Ethereum: A Secure Decentralised Generalised Transaction Ledger

### 2014-04-15 Paper.tex
Tweaks.

\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger}
\author{
    Dr. Gavin Wood\\
    CTO, Ethereum Project\\
    gavin@ethereum.org
}
\begin{document}

@@

In order to incentivise computation within the network, there needs to be an agreed method for transmitting value. To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH}. The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei. One Ether is defined as being $10^{18}$ Wei. There exist other subdenominations of Ether:

\par\bigskip
\begin{center}
\begin{tabular}{rl}
\toprule
Multiplier & Name \\
\midrule
$10^0$ & Wei \\
$10^{12}$ & Szabo \\
$10^{15}$ & Finney \\
$10^{18}$ & Ether \\
\bottomrule
\end{tabular}
\end{center}
\par

@@

The canonical difficulty of a block of header $H$, is defined as $D(H)$:
\begin{equation}
D(H) \equiv \begin{cases}
2^{22} & \quad \text{if genesis block}\\
{P(H)_H}_d + \frac{{P(H)_H}_d}{1024} & \quad \text{if $H_t < {P(H)_H}_t + 42$}\\
{P(H)_H}_d - \frac{{P(H)_H}_d}{1024} & \quad \text{otherwise}\\
\end{cases}
\end{equation}

@@

We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
\begin{equation}
g_0 \equiv \begin{cases}
  (|T_\mathbf{i}| + |T_\mathbf{b}|) G_{data} + G_{create} & \text{if} \quad T_r = 0 \\
  |T_\mathbf{d}| G_{data} + G_{call} & \text{otherwise} \\
\end{cases}
\end{equation}

where $|T_\mathbf{d}|$, $|T_\mathbf{i}|$ and $|T_\mathbf{b}|$ are the sizes, in bytes, of the transaction's associated data, initialisation and body EVM-code, respectively and $G$ is defined in Appendix \ref{app:fees}.

@@

We define the rollback state $\mathbb{S}_0$:
\begin{eqnarray}
\mathbb{S}_0 & \equiv & \mathbb{S} \quad \text{except:} \\
\mathbb{S}_0[S(T)]_{balance} & \equiv & \mathbb{S}[S(T)]_{balance} - c \\
\mathbb{S}_0[S(T)]_{nonce} & \equiv & \mathbb{S}[S(T)]_{nonce} + 1
\end{eqnarray}

@@

After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate. The Ether for the gas that was actually used is given to the miner, whose address is specified as the coinbase of the present block $B$. So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$:
\begin{eqnarray}
\mathbb{S}' & \equiv & \mathbb{S}_P \quad \text{except} \\
\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g' T_p \\
\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_P[m]_{balance} + (g - g') T_p \\
m & \equiv & {B_H}_b
\end{eqnarray}

In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
\begin{eqnarray}
\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g
\end{eqnarray}

@@

There are number of intrinsic parameters used when creating a contract: sender ($s$), nonce ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with the two arbitrary length byte arrays, $\mathbf{i}$ for the initialisation EVM code and $\mathbf{b}$ for the EVM code of the body.

We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\mathbb{S}_0$ to the tuple containing the new state, additional database entries and remaining gas $(\mathbb{S}_P, g')$, as in \ref{ch:transactions}:
\begin{equation}
(\mathbb{S}_P, g') \equiv \Lambda(\mathbb{S}_0, s, o, g, p, v, \mathbf{i}, \mathbf{b})
\end{equation}

@@

The exceptional halting function $Z$ is defined as:
\begin{equation}
Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \mathbf{\delta}_w = \varnothing \vee \lVert\mathbb{M}_S\rVert < \mathbf{\delta}_w
\end{equation}

@@

Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
\begin{eqnarray}
O(\mathbb{S}, \mathbb{M}, I) & \equiv & (\mathbb{S}', \mathbb{M}', I) \\
\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
\lVert\mathbb{M}'_{T}\rVert & \equiv & \lVert\mathbb{M}_{T}\rVert + \Delta \\
\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_T\rVert): \mathbb{M}'_{T_x} & \equiv & \mathbb{M}_{T_{x+\Delta}}
\end{eqnarray}

@@

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$.} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
&&&& $\mathbb{M}'_R = []$ \\
\midrule
0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] + \mathbb{M}_S[1]$ \\
\midrule
0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] \times \mathbb{M}_S[1]$ \\
\midrule
0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] - \mathbb{M}_S[1]$ \\
\midrule
0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
\midrule
0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
\midrule
0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] ^ {\mathbb{M}_S[1] }$ \\
\midrule
0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
&&&& $\mathbb{M}'_T[0] \equiv -\mathbb{M}_S[0]$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] < \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] > \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
\multicolumn{5}{l}{$\mathbb{M}_S[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_S[0]$} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \wedge \mathbb{M}_S[1]_i$ \\
\midrule
0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \vee \mathbb{M}_S[1]_i$ \\
\midrule
0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \oplus \mathbb{M}_S[1]_i$ \\
\midrule
0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
&&&& $\forall i \in [0..7]: \mathbb{M}'_T[0]_i \equiv \begin{cases} \mathbb{M}_S[1]_{(i + 8\mathbb{M}_S[0])} & \text{if} \quad \mathbb{M}_S[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathtt{\tiny SHA3}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1]) \div 32 })$ \\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{30s: Environmental Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
&&&& $\mathbb{M}'_T[0] \equiv I_a$ \\
\midrule
0x31 & {\small BALANCE} & 0 & 1 & Get balance of currently executing account. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}[I_a]_{balance}$ \\
\midrule
0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
&&&& $\mathbb{M}'_T[0] \equiv I_o$ \\
&&&& This is the sender of original transaction; it's never a contract. \\
\midrule
0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
&&&& $\mathbb{M}'_T[0] \equiv I_c$ \\
&&&& This is the address of the account that is directly responsible for this execution. \\
\midrule
0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
&&&& $\mathbb{M}'_T[0] \equiv I_v$ \\
\midrule
0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 32) ]$ \\
&&&& This pertains to the input data passed with the message call instruction or transaction. \\
\midrule
0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
&&&& This pertains to the input data passed with the message call instruction or transaction. \\
\midrule
0x37 & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv I_p$ \\
&&&& This is gas price specified by the originating transaction.\\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
&&&& $\mathbb{M}'_T[0] \equiv {I_H}_p$ \\
&&&& $I_{H_p}$ is the previous block's hash. \\
\midrule
0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
&&&& $\mathbb{M}'_T[0] \equiv {I_H}_b$ \\
\midrule
0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
&&&& $\mathbb{M}'_T[0] \equiv {I_H}_t$ \\
\midrule
0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
&&&& $\mathbb{M}'_T[0] \equiv n({I_H})$ \\
&&&& $n(I_H)$ is the total number of ancestor blocks of the current block. \\
\midrule
0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
&&&& $\mathbb{M}'_T[0] \equiv {I_H}_d$ \\
\midrule
0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
&&&& $\mathbb{M}'_T[0] \equiv 10^6$ \\
&&&& Presently, the gas limit is hard-coded in the protocol. \\
\bottomrule
\end{tabular*}

>>>>

\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger}
\author{
    Dr. Gavin Wood\\
    Co-Founder \& CTO, Ethereum Project\\
    gavin@ethereum.org
}
\begin{document}

@@

In order to incentivise computation within the network, there needs to be an agreed method for transmitting value. To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH}. The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei. One Ether is defined as being $10^{18}$ Wei. There exist other subdenominations of Ether:
\par`
\begin{center}
\begin{tabular}{rl}
\toprule
Multiplier & Name \\
\midrule
$10^0$ & Wei \\
$10^{12}$ & Szabo \\
$10^{15}$ & Finney \\
$10^{18}$ & Ether \\
\bottomrule
\end{tabular}
\end{center}
\par

@@

The canonical difficulty of a block of header $H$, is defined as $D(H)$:
\begin{equation}
D(H) \equiv \begin{cases}
2^{22} & \quad \text{if genesis block}\\
{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \quad \text{if $H_t < {P(H)_H}_t + 42$}\\
{P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \quad \text{otherwise}\\
\end{cases}
\end{equation}

@@

We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
\begin{equation}
g_0 \equiv \begin{cases}
g_d + G_{create} & \text{if} \quad T_r = 0 \\
g_d + G_{call} & \text{otherwise} \\
\end{cases}
\end{equation}

where

\begin{equation}
g_d \equiv \begin{cases} (\lVert T_\mathbf{i}\rVert + \lVert T_\mathbf{b}\rVert) G_{data} & \text{if} \quad T_r = 0 \\
 \lVert T_\mathbf{d}\rVert G_{data} & \text{otherwise}
 \end{cases}
\end{equation}

where $\lVert T_\mathbf{d}\rVert$, $\lVert T_\mathbf{i}\rVert$ and $\lVert T_\mathbf{b}\rVert$ are the sizes, in bytes, of the transaction's associated data, initialisation and body EVM-code, respectively and $G$ is defined in Appendix \ref{app:fees}. The portion of $g_0$ that will not be paid to the miner is $g_d$, and is removed from the system.

% Explain g_d reason?

@@

We define the rollback state $\mathbb{S}_0$:
\begin{eqnarray}
\mathbb{S}_0 & \equiv & \mathbb{S} \quad \text{except:} \\
\mathbb{S}_0[S(T)]_{balance} & \equiv & \mathbb{S}[S(T)]_{balance} - v_0 \\
\mathbb{S}_0[S(T)]_{nonce} & \equiv & \mathbb{S}[S(T)]_{nonce} + 1
\end{eqnarray}

@@

After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate. The Ether for the gas that was actually used is given to the miner, whose address is specified as the coinbase of the present block $B$. So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$:
\begin{eqnarray}
\mathbb{S}' & \equiv & \mathbb{S}_P \quad \text{except} \\
\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g' T_p \\
\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_P[m]_{balance} + (T_g - g' - g_d) T_p \\
m & \equiv & {B_H}_b
\end{eqnarray}

%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
%\begin{eqnarray}
%\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g
%\end{eqnarray}

@@

There are number of intrinsic parameters used when creating a contract: sender ($s$), nonce ($n$), available gas ($g$), gas price ($p$), endowment ($v$) together with the two arbitrary length byte arrays, $\mathbf{i}$ for the initialisation EVM code and $\mathbf{b}$ for the EVM code of the body.

We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\mathbb{S}_0$ to the tuple containing the new state, additional database entries and remaining gas $(\mathbb{S}_P, g')$, as in \ref{ch:transactions}:
\begin{equation}
(\mathbb{S}_P, g') \equiv \Lambda(\mathbb{S}_0, s, n, g, p, v, \mathbf{i}, \mathbf{b})
\end{equation}

@@

The exceptional halting function $Z$ is defined as:
\begin{equation}
Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}) \vee \mathbf{\delta}_w = \varnothing \vee \lVert\mathbb{M}_S\rVert < \mathbf{\delta}_w
\end{equation}

@@

Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
\begin{eqnarray}
O(\mathbb{S}, \mathbb{M}, I) & \equiv & (\mathbb{S}', \mathbb{M}', I) \\
\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
\lVert\mathbb{M}'_{S}\rVert & \equiv & \lVert\mathbb{M}_{T}\rVert + \Delta \\
\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_S\rVert): \mathbb{M}'_{T_x} & \equiv & \mathbb{M}_{T_{x+\Delta}}
\end{eqnarray}

@@

\section{Acknowledgements}

Vitalik Buterin first proposed in the original Ethereum whitepaper (first distributed in December 2013) moving from a single state-transaction function with trivial state to something much more general and expansive. Useful corrections and suggestions were provided by a number of others from the Ethereum community including Aeron Buchanan

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$.} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
&&&& $\mathbb{M}'_R = []$ \\
\midrule
0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] + \mathbb{M}_S[1]$ \\
\midrule
0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] \times \mathbb{M}_S[1]$ \\
\midrule
0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] - \mathbb{M}_S[1]$ \\
\midrule
0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
\midrule
0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
\midrule
0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] ^ {\mathbb{M}_S[1] }$ \\
\midrule
0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
&&&& $\mathbb{M}'_S[0] \equiv -\mathbb{M}_S[0]$ \\
&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
\midrule
0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] < \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] > \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
\midrule
0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
\multicolumn{5}{l}{$\mathbb{M}_S[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_S[0]$} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_S[0]_i \equiv \mathbb{M}_S[0]_i \wedge \mathbb{M}_S[1]_i$ \\
\midrule
0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_S[0]_i \equiv \mathbb{M}_S[0]_i \vee \mathbb{M}_S[1]_i$ \\
\midrule
0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
&&&& $\forall i \in [0..255]: \mathbb{M}'_S[0]_i \equiv \mathbb{M}_S[0]_i \oplus \mathbb{M}_S[1]_i$ \\
\midrule
0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
&&&& $\forall i \in [0..7]: \mathbb{M}'_S[0]_i \equiv \begin{cases} \mathbb{M}_S[1]_{(i + 8\mathbb{M}_S[0])} & \text{if} \quad \mathbb{M}_S[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathtt{\tiny SHA3}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1]) \div 32 })$ \\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{30s: Environmental Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
&&&& $\mathbb{M}'_S[0] \equiv I_a$ \\
\midrule
0x31 & {\small BALANCE} & 0 & 1 & Get balance of currently executing account. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{S}[I_a]_{balance}$ \\
\midrule
0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
&&&& $\mathbb{M}'_S[0] \equiv I_o$ \\
&&&& This is the sender of original transaction; it is never a contract. \\
\midrule
0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
&&&& $\mathbb{M}'_S[0] \equiv I_c$ \\
&&&& This is the address of the account that is directly responsible for this execution. \\
\midrule
0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
&&&& $\mathbb{M}'_S[0] \equiv I_v$ \\
\midrule
0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
&&&& $\mathbb{M}'_S[0] \equiv I_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
&&&& This pertains to the input data passed with the message call instruction or transaction. \\
\midrule
0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
&&&& $\mathbb{M}'_S[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
&&&& This pertains to the input data passed with the message call instruction or transaction. \\
\midrule
0x37 & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
&&&& $\mathbb{M}'_S[0] \equiv I_p$ \\
&&&& This is gas price specified by the originating transaction.\\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
&&&& $\mathbb{M}'_S[0] \equiv {I_H}_p$ \\
&&&& $I_{H_p}$ is the previous block's hash. \\
\midrule
0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
&&&& $\mathbb{M}'_S[0] \equiv {I_H}_b$ \\
\midrule
0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
&&&& $\mathbb{M}'_S[0] \equiv {I_H}_t$ \\
\midrule
0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
&&&& $\mathbb{M}'_S[0] \equiv n({I_H})$ \\
&&&& $n(I_H)$ is the total number of ancestor blocks of the current block. \\
\midrule
0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
&&&& $\mathbb{M}'_S[0] \equiv {I_H}_d$ \\
\midrule
0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
&&&& $\mathbb{M}'_S[0] \equiv 10^6$ \\
&&&& Presently, the gas limit is hard-coded in the protocol. \\
\bottomrule
\end{tabular*}
